---
title: "Lec4-5 In-Class Exercise"
author: "Luc"
date: "2023-10-04"
output: pdf_document
---

```{r}
d <- read_csv("/Users/lucmekouar/Desktop/POLI 3148/!!POLI_3148_repo/data-1984_2022/vdem_1984_2022_external.csv")

```
1. Coodbook lookup. Look up the codebook, answer the following questions:
i.    What indicators regarding the quality of education are available in the V-Dem datasets?
 - Educational equality (v2peedueq)
 - Educational inequality, Gini (e_peedgini)
 

```{r}
d |> select(country_name) |> distinct()
  
```




ii.   What are the data’s coverage (i.e., for which countries and years do we have data?)

```{r}
d |> select(country_name) |> distinct() 
```

```{r}
d |> select(year) |> distinct()
```



iii.  What are their sources? Provide the link to least 1 source.
https://www.v-dem.net/data/dataset-archive/ ??




2. Subset by columns
i.    Create a dataset containing only the country-year identifiers and indicators of education quality.

ii.   Rename the columns of education quality to make them informative.

3. Subset by rows
i.    List 5 countries-years that have the highest education level among its population.

ii.   List 5 countries-years that suffer from the most severe inequality in education.

4. Summarize the data
i.    Check data availability: For which countries and years are the indicators of education quality available?

ii.   Create two types of country-level indicators of education quality
a.        Average level of education quality from 1984 to 2022

b.        Change of education quality from 1984 to 2022

iii.  Examine the data and briefly discuss: Which countries perform the best and the worst in terms of education quality in the past four decades?


